---
title: "8-shig_luminex_serimmune_comp"
author: "Niko Walas"
date: "2025-01-23"
format:
  html:
    self-contained: true
    toc: true
    embed-resources: true
    code-fold: true
    link-external-newwindow: true
execute: 
  echo: true
  warning: false
  message: false
---

## File overview

The purpose of this analysis, is to compare serimmune summary measures for Shigella, with those from the Serimmune and TAC assays (IgG and PCR). Important to note:

1.  The TAC-PCR results for Shigella are only available for visit 2.

The following comparisons are made below

1) Agreement between PCR status and IgG status

2) PCR status and IWAS summary measures
    a) meanIWAS
    b) maxIWAS
    c) Number of epitope hits

We conclude:

1.  There is alignment between PCR status and IgG seropositivity. 89% of PCR+ samples seroconverted. A large proportion of seropositive samples did not have a corresponding PCR+ result. 


```{r preamble, echo = FALSE, message = FALSE}
#----------------------------------
# load the config file
#----------------------------------
library(here)
source("0-shig-config.r")
library(epiR)
pcols <- brewer.pal(n=3,"Dark2")[c(3,2)]
visitcols <- c("#E69F00", "#56B4E9", "#009E73")
seroconvcols <- c("orange", "gold", "darkblue", "grey", "red")
  
#----------------------------------
# load the protein level summary
#----------------------------------
dp1 <- read_csv(here("data","shigella_ipa_sum_prot.csv")) %>% select(-1) %>% rename(no_epitopes = no_pos_ep)

#----------------------------------
# load the pathogen level summary
#----------------------------------
dpath1 <- read_csv(here("data","shigella_ipa_sum_path.csv")) %>% select(-1) %>% 
  rename(no_epitopes = epitope_count, 
         no_proteins = protein_count, 
         sampleid = SampleID)

#----------------------------------
# load metadata
#----------------------------------
metadata <- read_csv(here("data/washb-serimmune-metadata-230616.csv")) %>% pivot_longer(cols = 2:4, names_to = "timepoint", values_to = "SampleID")
metadata2 <- metadata %>% select(SampleID, arm, sex, birthord_cat, momage, agedays_base, agedays_mid, agedays_end) %>% pivot_longer(6:8, names_to = "age_visit") %>% 
  filter(str_detect(SampleID, "B1") & age_visit == "agedays_base" | 
         str_detect(SampleID, "M1") & age_visit == "agedays_mid" |
         str_detect(SampleID, "E1") & age_visit == "agedays_end") %>% 
  rename(sampleid = SampleID, age_days = value)

#----------------------------------
# load TAC-PCR
# This file was generated for serimmune in the sample-selection folder on box
# /Users/nwalas/Library/CloudStorage/Box-Box/washb-serimmune/sample-selection/rmd/washb-serimmune-sample-selection-v4.0
# Tac data was only run at visit two
#----------------------------------
shig_tac <- read_csv(here("data/washb-serimmune-metadata-230616.csv")) %>% select(sampleid_mid, arm, agedays_mid, Shig_EIEC) %>% 
  rename(sampleid = sampleid_mid) %>% 
  mutate(shig_tac = ifelse(Shig_EIEC == "Yes", "PCR+", "PCR-")) %>% select(-Shig_EIEC)

```

DF creation

```{r}
#----------------------------------
# create a merged data-frames for comparisons below
#---------------------------------
df <- dpath1 %>% full_join(shig_tac, by = "sampleid")
# add in age in day
df <- df %>% full_join(metadata2 %>% select(sampleid, age_visit, age_days), 
                       by = c("sampleid")) %>% 
  select(-agedays_mid) %>% # reduce redundancy
  arrange(sampleid, visit) %>% 
  mutate(sero = ifelse(sero_pos_final == "POS", "IgG+", "IgG-"))

# need to fill the NA vaules for pcr with the value from the visit 2 for the plot
df <- df %>% group_by(child) %>% 
  mutate(shig_tac2 = coalesce(shig_tac, first(na.omit(shig_tac))), 
         sero_visit2 = sero[time_point == 2][1], 
         arm2 = coalesce(arm, first(na.omit(arm))))

# assign levels 
p_df <- df %>% group_by(child) %>% 
  mutate(serimmune_pos_afterv2 = any(sero == "IgG+" & visit %in% c("Visit 2", "Visit 3"), na.rm = TRUE), 
         serimmune_pos_afterv2 = ifelse(serimmune_pos_afterv2 == "TRUE", "IgG+", "IgG-"))

p_df <- p_df %>% 
  mutate(arm2 = factor(arm2, levels = c("Control", "Nutrition + WSH")),
         sero = factor(sero, levels = c("IgG+", "IgG-")), 
         serimmune_pos_afterv2 = factor(serimmune_pos_afterv2, levels= c("IgG+", "IgG-")),
        shig_tac = factor(shig_tac, levels = c("PCR+", "PCR-")), 
        shig_tac2 = factor(shig_tac2, levels = c("PCR+", "PCR-"))
        )

# add in timepoint seroconversions for the 3 IgG measures
p_df <- p_df %>% group_by(child) %>%
  arrange(child,time_point) %>%
  mutate(serimmune_seropos_lead = lead(sero),
         seroconv_serimmune = case_when(
           serimmune_seropos_lead  == "IgG+" & sero == "IgG-" ~ "Seroconversion",
           serimmune_seropos_lead  == "IgG-" & sero == "IgG+" ~ "Seroreversion",
           serimmune_seropos_lead  == "IgG+" & sero== "IgG+" ~ "Seropositive",
           serimmune_seropos_lead  == "IgG-" & sero == "IgG-" ~ "Seronegative"
        
           )
         
         )
         
# assign levels 
p_df <- p_df %>% mutate(
  seroconv_serimmune = factor(seroconv_serimmune, 
                              levels = c("Seropositive","Seroconversion", 
                                         "Seroreversion", "Seronegative", NA))

)
```

# PCR vs Serimmune

## 1) PCR vs Serimmune IWAS

### a) Table of agreement

```{r}

t1<- table(p_df$shig_tac, p_df$serimmune_pos_afterv2)

cat("Shigella PCR vs Serimmune IWAS, visit 2\n\n")
t1 

# Kappa calculation 
k_out <- epiR::epi.kappa(t1)
k_est <- k_out$kappa$est
kappa_lci  <- k_out$kappa$lower
kappa_uci  <- k_out$kappa$upper

# Agreement
agree_est <- k_out$prop.agree$obs
agree_ci <- binom.test(sum(diag(t1)), sum(t1))$conf.int
agree_lci <- agree_ci[1]
agree_uci <- agree_ci[2]

agreement <- paste0(round(agree_est*100, 2), " (", round(agree_lci*100, 2), ", ", round(agree_uci*100, 2), ")")
kappa <-  paste0(round(k_est, 2), " (", round(kappa_lci, 2), ", ", round(kappa_uci, 2), ")")

a_k <- data.frame(agreement, kappa) %>% pivot_longer(1:2, names_to = " ")
cat("Percent agreement and kappa- PCR and Serimmune IgG, visit 2\n\n")
a_k

pcr_iwas_agreement <- a_k 
```
### b) Number epitope hits (all Ipa proteins)

```{r}
## summary of epitope hits by age
#| label: plot epitope hits for shigella 

# Plot 1: Visit 2 ONLY 
plot_noep_shigella_v2 <- ggplot(data=p_df %>% filter(visit == "Visit 2"), 
                                aes(x = no_epitopes, 
                                    color=factor(arm2), 
                                    fill = factor(arm2))) + 
  geom_bar(alpha = 0.5, position = "dodge") +
  facet_grid(serimmune_pos_afterv2~ ~ shig_tac2) +
  scale_color_manual(values=pcols) +
  scale_fill_manual(values=pcols) +
  labs(x = "Number of epitope hits",y = "Number of children", 
       title = " Individual child epitope hits",
       subtitle = "by PCR infection status (visit 2) and seropositivity status (visit 2 and3)", 
       color = NULL, 
       fill = NULL) + theme_classic() + 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6))

plot_noep_shigella_v2

# Plot 2: Visit 2, 3 (to exclude maternal antibodies)
plot_noep_shigella_v23 <- ggplot(data=p_df %>% filter(visit != "Visit 1"), aes(x=age_days, y = no_epitopes, 
                                             group=child, 
                                             color=factor(seroconv_serimmune))) + 
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = 2, linetype = "dashed", color = "darkgrey") +
  facet_grid(serimmune_pos_afterv2~ ~ shig_tac2) +
  #geom_point() + 
  scale_color_manual(values=seroconvcols) +
  labs(x = "Age (days)",y = "Number of epitope hits", 
       title = " Individual child epitope hits",
       subtitle = "by PCR infection status (visit 2) and seropositivity status (visit 2 and 3)", 
       color = NULL) + theme_classic() + 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6)) +
  coord_cartesian(ylim = c(0, 15))


plot_noep_shigella_v23
ggsave("../figures/pcr_iwas_agreement_epitopehits.png", plot_noep_shigella_v23, width = 8, height =5)
```


### c) meanIWAS 

Important to note: the assignment of PCR/Serimmune(IgG) agreement is assigned only at the visit 2 timepoint
This is the meanIWAS across all five Ipa proteins used in the seropositivity assignment

```{r}
## summary of meanIWAS by age
#| label: plot meanIWAS for shigella 
         
# Plot 1: Visit 2 ONLY
plot_meaniwas_shigella_v2 <- ggplot(data=p_df %>% filter(visit == "Visit 2"), 
                                aes(x = mean_IWAS, 
                                    color=factor(arm2), 
                                    fill = factor(arm2))) + 
  geom_histogram(alpha = 0.5, position = "dodge") +
  facet_grid(serimmune_pos_afterv2 ~ shig_tac2) +
  #geom_point() + 
  scale_color_manual(values=pcols) +
  scale_fill_manual(values=pcols) +
  labs(x = "meanIWAS",y = "Number of children", 
       title = " Individual child meanIWAS",
       subtitle = "by PCR infection status (visit 2) and seropositivity status (visit 2 and 3)",  
       color = NULL, 
       fill = NULL) + theme_classic() + 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6))

plot_meaniwas_shigella_v2


# Plot 2: Visits 2, 3 (to exclude maternal contribution)
plot_meaniwas_shigella_v23 <- ggplot(data=p_df %>% filter(visit != "Visit 1"), 
                                     aes(x=age_days, y = mean_IWAS, 
                                             group=child, 
                                             color=seroconv_serimmune)) + 
  geom_line(alpha = 0.5) +
  facet_grid(serimmune_pos_afterv2 ~ shig_tac2) +
  scale_color_manual(values=seroconvcols) +
  labs(x = "Age (days)",y = "meanIWAS", 
       title = " Individual child meanIWAS",
       subtitle = "by PCR infection status (visit 2) and seropositivity status (visit 2 and 3)", 
       color = NULL) + theme_classic() + 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6))


plot_meaniwas_shigella_v23
ggsave("../figures/meaniwas_pcr_visit2.png", plot_meaniwas_shigella_v23,  width = 8, height =5)
```

### d) maxIWAS (across all Ipa proteins)

```{r}
## summary of maxIWAS by age
#| label: plot maxIWAS for shigella

# Plot 1: Visits 2, 3 (to exclude maternal contribution)
plot_maxiwas_shigella_v23 <- ggplot(data=p_df %>% filter(visit != "Visit 1"), 
                                    aes(x=age_days, y = max_IWAS, 
                                             group=child, 
                                             color=factor(seroconv_serimmune))) + 
  geom_line(alpha = 0.5) +
  facet_grid(serimmune_pos_afterv2 ~ shig_tac2) +
  #geom_point() + 
  scale_color_manual(values=seroconvcols) +
  labs(x = "Age (days)",y = "maxIWAS", 
       title = " Individual child maxIWAS",
       subtitle = "by PCR infection status (visit 2) and seropositivity status (visit 2 and 3)", 
       color = NULL) + theme_classic() + 
  theme(panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6))



plot_maxiwas_shigella_v23
ggsave("../figures/maxiwas_pcr_visit2.png", plot_maxiwas_shigella_v23,  width = 8, height =5)

```



