---
title: "9-shig_validation_motif"
output: html_document
date: "2026-04-10"
---

File description: 

The purpose of this file was to validate the motif findings in our samples with known cases of Shigella discovered by Richelle Charles group. 

Note: only Ipa protein motifs were used in this analysis

```{r preamble, echo = FALSE, message = FALSE}
#----------------------------------
# load the config file
#----------------------------------
library(here)
source("0-shig-config.r")
library(epiR)
library(openxlsx)
pcols <- brewer.pal(n=3,"Dark2")[c(3,2)]
visitcols <- c("#E69F00", "#56B4E9", "#009E73")
seroconvcols <- c("orange", "gold", "darkblue", "grey", "red")
  
#----------------------------------
# load the motif cleaned summary
#----------------------------------
df_motifs <- read_csv(here("data/shigella_motif_clean.csv")) 

#----------------------------------
# load the pathogen level summary
#----------------------------------
dpath1 <- read_csv(here("data","shigella_ipa_sum_path.csv")) %>% select(-1) %>% 
  rename(no_epitopes = epitope_count, 
         no_proteins = protein_count, 
         sampleid = SampleID)


#----------------------------------
# load metadata
#----------------------------------
metadata <- read_csv(here(metadatadir,"washb-serimmune-metadata-230616.csv")) %>% pivot_longer(cols = 2:4, names_to = "timepoint", values_to = "SampleID")
metadata2 <- metadata %>% select(SampleID, arm, sex, birthord_cat, momage, agedays_base, agedays_mid, agedays_end) %>% pivot_longer(6:8, names_to = "age_visit") %>% 
  filter(str_detect(SampleID, "B1") & age_visit == "agedays_base" | 
         str_detect(SampleID, "M1") & age_visit == "agedays_mid" |
         str_detect(SampleID, "E1") & age_visit == "agedays_end") %>% 
  rename(sampleid = SampleID, age_days = value)

#----------------------------------
# load sera scores 
#----------------------------------
shig_sera <- read_csv(here("data/shigella_sera_score.csv"))

#----------------------------------
# load known positive shigella data 
#----------------------------------
motif_knownpos <- read_excel(here("data/shigella_epitope_analysis_final.xlsx"), skip = 5 )
```

## Polish the known positive sample set 
```{r}
# Need a lot of polishing
# 5) rename sensitivity
motif_knownpos <- motif_knownpos[, 1:8] 
motif_knownpos <- motif_knownpos %>%
  rename(shigella_cohort_sensitivity = `Shigella Cohort Discovery Sensitivity`, 
         day_2_sensitivity = D2, 
         day_7_sensitivity = D7, 
         day_30_sensitivity = D30,
         example_family_motif = `Example Family Motif`, 
         family = Family, 
         antigen = Antigen, 
         epitope = Epitope)

motif_knownpos$shigella_cohort_sensitivity <- as.numeric(motif_knownpos$shigella_cohort_sensitivity)

motif_knownpos <- motif_knownpos %>% arrange(desc(shigella_cohort_sensitivity)) 

# filter to Ipa motif
motif_knownpos <- motif_knownpos %>% filter(str_detect(antigen, "Ipa"))

# 23 Ipa specific motifs
```

## Investigate motifs from IWAS seropositivity

```{r}
# merge the IWAS seropositivity calls 
df_motifs <- df_motifs %>% left_join(dpath1 %>% select(sampleid, sero_pos_final), by = c("sampleid"))

# summarise the number of positive motifs based on our IWAS seropositivity calls
motif_summary <- df_motifs %>%
  group_by(sampleid) %>%
  summarise(
    any_motif_pos = ifelse(any(motif_pos == "POS", na.rm = TRUE), "any_motif+", "any_motif-"),
    motif_pos_count = sum(motif_pos == "POS", na.rm = TRUE),
    sero_pos_final = first(sero_pos_final)
  ) %>% mutate(sero_pos_final = ifelse(sero_pos_final == "POS", "IWAS IgG+", "IWAS IgG-"))


# table of agreement 
table(motif_summary$any_motif_pos, motif_summary$sero_pos_final)

# plot of number of motifs positive by IWAS assignment
ggplot(motif_summary %>% filter(sero_pos_final != 0), aes(x = motif_pos_count)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  facet_wrap(~ sero_pos_final) +
  labs(
    x = "Motif Positive Count",
    y = "Number of Samples",
    title = "Distribution of Motif Counts by Serostatus"
  ) +
  theme_minimal()
```

## Look at top motif hits from known positives
```{r}
# top 5 epitope hits
top5 <- motif_knownpos %>%
  slice_max(order_by = shigella_cohort_sensitivity, n = 5)
top5

# how many motifs for ipa proteins ony

range(motif_knownpos$shigella_cohort_sensitivity)
# some of these motifs have low sensitivity for shigella ?
```

## Overlap between detected motifs
```{r}

# Number of shared motifs
shared_motifs <- intersect(motif_knownpos$example_family_motif, df_motifs$epitope_motif)
shared_motifs

# difference between the motifs detected in both
unique_to_ourdata <- setdiff(df_motifs$epitope_motif, motif_knownpos$example_family_motif)
unique_to_ourdata

missed_from_ourdata <- setdiff(motif_knownpos$example_family_motif, df_motifs$epitope_motif)
missed_from_ourdata
```
Remaining questions: 
1. How was sensitivity determined? Statistical process for what was considered a motif and a hit?
2. How to interpret low sensitivity motifs that are present
3. How to interpret motifs present in both IWAS positive and IWAS negative samples

# Alignment of IWAS categorization for each Ipa motif, and corresponding sensitivity
```{r}
t1 <- df_motifs %>% group_by(sero_pos_final, probable_shigella_antigen) %>% 
  mutate(sero_pos_final = ifelse(sero_pos_final == "POS", "IWAS IgG+", "IWAS IgG-")) %>% 
  filter(motif_pos == "POS") %>% 
  count(epitope_motif) %>%
  pivot_wider(
    names_from = sero_pos_final,
    values_from = n,
    values_fill = 0
  ) %>% 
  mutate(n_with_motif = `IWAS IgG+` + `IWAS IgG-`, 
         percent_motif_IWASpos = (`IWAS IgG+`/n_with_motif)*100) %>% 
  left_join(motif_knownpos %>% select(example_family_motif, shigella_cohort_sensitivity, 
                                      day_2_sensitivity, day_7_sensitivity, 
                                      day_30_sensitivity) %>% 
              mutate(epitope_motif = example_family_motif) %>% select(-example_family_motif), 
            by = "epitope_motif") %>% 
  mutate(antigen = case_when(str_detect(probable_shigella_antigen, "IpaA") ~ "IpaA", 
                             str_detect(probable_shigella_antigen, "IpaB") ~ "IpaB",
                              str_detect(probable_shigella_antigen, "IpaC") ~ "IpaC",
                             str_detect(probable_shigella_antigen, "IpaD") ~ "IpaD",
                             TRUE ~ "CHECK")) %>% 
  ungroup() %>% 
  select(-probable_shigella_antigen) %>% 
  select(antigen, epitope_motif, `IWAS IgG+`, `IWAS IgG-`, n_with_motif, 
         percent_motif_IWASpos, shigella_cohort_sensitivity, 
         day_2_sensitivity, day_7_sensitivity, day_30_sensitivity)


# table of motif sensitivity in our dataset
kable(t1)
```

## How can I assess the alignment between our results and the known positive results? 
```{r}
names(t1)

supp_tab6 <- t1 %>%select(antigen, epitope_motif, n_with_motif, percent_motif_IWASpos, day_30_sensitivity) %>%
  mutate(prop_withmotif = round(n_with_motif/360*100, 2)) %>% 
  select(-n_with_motif) %>% 
  select(antigen, epitope_motif, prop_withmotif, percent_motif_IWASpos, day_30_sensitivity) %>% 
           arrange(desc(day_30_sensitivity)) %>% 
  mutate(day_30_sensitivity = round(day_30_sensitivity*100, 2), 
         percent_motif_IWASpos = round(percent_motif_IWASpos, 2))

supp_tab6

write.xlsx(supp_tab6, "../output/supptable6.xlsx")
```













